---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Download file from the web and unzip it.
```{r echo=TRUE, cache=TRUE}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("data")) {
  dir.create("data")
  download.file(fileUrl, "./data/activity.zip")
  unzip("data/activity.zip", exdir = "data")
}
```
Read file and save it in activity variable.
```{r cache=TRUE}
activity <- read.csv("data/activity.csv")
```
Show information about in data set using str function
```{r}
str(activity)
```
Transform date to date type
```{r}
activity <- transform(activity, date = as.Date(date))
str(activity)
```

## What is mean total number of steps taken per day?
Calculate the total number of steps per day
```{r}
totalStepsPerDay <- sapply(split(activity$steps, activity$date), sum)
```
Sapply function returns a vector, so for plots we should convert it to data frame
```{r}
totalStepsPerDay <- as.data.frame(totalStepsPerDay)
```
Make a histogram of the total number of steps taken each day
1. Load package ggplot2
```{r}
library(ggplot2)
```
2. Create the histogram
```{r}
p <- ggplot(data = totalStepsPerDay, aes(x = totalStepsPerDay)) + geom_histogram(fill = "green", binwidth = 1000) + labs(x = "Steps", y = "Frequency", title = "Histogram of total steps per day")
p
```

Calculate the mean of the total number of steps taken per day
```{r}
mean(totalStepsPerDay$totalStepsPerDay, na.rm = TRUE)
```
Calculate the median of the total number of steps taken per day
```{r}
median(totalStepsPerDay$totalStepsPerDay, na.rm = TRUE)
```
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
